---
title: "Rmd workshop 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE}

set.seed(67)

```

# {.tabset}

In this weeks excercise we are going to go over how to simulate coin flips
in R to build an example of basic simulation of probabilities and then build off
and apply this understanding using a real-world scenario.  We will interpret this real 
world scenario using probabilistic reasoning to determine whether our real world outcome
is to be expected or fair.<br />

*Don't forget the tabs just below this setence when you are ready to move
on to the next section.*



## **Flipping a Coin**

To begin learning probability simulation in R we will simulate a coin flip. 
In another post I will do more number games and simulations. I encourage you 
to always try to follow along in R on your own computer.

<br />  

**1.** Create a character vector containing "H" and "T", store in object 'x'. <br /> 

```{r}

# 'H' will represent heads and 'T' will represent tails.

x <- c("H", "T")

````
<br />



**2.** Pass object x through the sample() function, specfying the functions
  arguments such that we sample our original vector 'x' 10 times, with
  replacement, and store this sample in the object 'y'. <br />  

```{r}
# Sample 'x' 10 times with replacement:

y <- sample(x, size = 10, replace = TRUE)

```
  
**Tip**: If you forget the arguments a function can work with or whether an argument
has a default value (meaning it is not mandatory for the user to specify its
value), you can look up the documentaion by typing ?function(). Or, an even
quicker way is to type args(function). <br />   


args(sample) returns: <br /> 
function (x, size, replace = FALSE, prob = NULL) <br />   


This means the sample function requires the user to specify what is being sampled (x)
and the size of the sample (size), but will by default sample this object or vector
without replacement (replace = ) and not specify any probability weightings (prob = NULL). <br />  

<br />


  
**3.** Print y to the console. <br /> 

```{r, eval= FALSE}
y

```

Below is our output for this sample. Remember if you're following along in R you will likely get a 
different outcome.

```{r, echo= FALSE}
y

````
<br />

In stead of counting the number of heads in our outcome (this would be difficult with a larger sample),
let's use a *logical operator* to see how many heads we got in our sample of 10 flips.<br />  


Read "==" as "exactly equal to." We are asking R to return a vector of logical
values which will read TRUE when a value within our sampling object (y) is exactly equal to
"H" and FALSE when when a value is not exactly equal to "H" (in this case, when the value
is "T").

```{r}
# Asking which values of y are exactly ewual to "H"

y == "H"

```
<br />

Did this initial step solve our problem of counting values in our output?
No. We should instead subset our object y to include only outcomes which were
heads and store this in a new object so we can apply useful functions to this object. 

We can then call length() on this new object to ask R to do the counting for us.
Because our new object was filled with only outcomes of y which were heads, 
if we ask for the length of the vector, it informs us how many heads we flipped.
The length function is very useful.


```{r}
# Subset y to include only heads, store as new object and call length

how_many <- y[ y == "H" ]
length(how_many)

```

```{r, echo = FALSE}
t <- length(how_many)
```
<br />


Because the probability of flipping a head or tail is 0.5, we would expect to see 5 heads
and 5 tails after flipping a coin 10 times, on average. In our example, our observed outcome
matched this expect outcome exactly. This won't always be the case. Remember, in 10 flips of a 
coin it is very unlikely, but possible, to get all heads. This reflects the concept of sampling
variation <br />  
<br />



**BONUS** 

Using the sample() function is certainly the best way to simulate this experiment,
but to add to our understanding by simulating the coin flip experiment another way
using the runif() function and a for loop.<br />

We can ask R to return a random number between 0 and 1, with equal probabilities,
with the *runif()* function. Notice I wrapped the round() function around runif(),
asking R to round the output of runif() to 2 decimal places just for readability
of this excercise. <br />

We know the probability of rolling a head or a tail is 0.50, so we could say that
any outcome of runif() less than .05 represents landing on a head, where any outcome
greater than .50 represents landing on a tail.

```{r}

# Get one random number between 0 and 1:

randNum <- round(runif(1), digits = 2)
randNum

```
<br />

Notice the runif() function asks for a value that specifies the amount of
times you want to generate a number between 0 and 1. We could change this
to 1000 to get 1000 random numbers. Let's try with 15.

```{r}
randNum <- round(runif(15), digits = 2)
randNum

```
<br />

Let us create a for loop in R with the following rules: <br />

a) Generate a random number between 1 and 10 (randNum) <br />
b) If randNum < .05, outcome = heads <br />
c) If randNum > .05, outcome = tails <br />
d) Repeat N times (number of flips) <br />
<br />

First define N as the number of coin flips to simuate (100).
```{r}
N <- 100
```
<br />

Call runif() on N to generate 100 random numbers between 0 and 1. Store in object randNum. 
```{r}
randNum <- runif(N)
```
<br />

Fill values into an empty matrix (notice the NA where we would usually
reference a data frame) matrix. We set the dimesions of the matrix to be one
columns and as many rows as our number of trials (N). 

This loop will simulate our experiment of flipping a coin 100 times.

```{r}

x <- matrix(NA,nrow = N,ncol = 1)
for (i in 1:N)
{
 if(randNum[i] < 0.5 )
 { x[i,1] = "H" } else 
 { x[i,1] = "T" }
}


```
<br />

Now let's take a look at our probability of heads and tails. We will introduce
a new peice of code here which uses logical operation to find only heads and only tails
within a function instead of saving these values to a new object like we did in the
first experiment.<br />

First, let's see the probability of getting heads:

```{r}


```
<br />

Next, let's see the probability of getting tails: 

```{r}

```
<br />



## Flip a coin many times

Now flip a coin 100,000 times and count how many heads you get.
We are going to use the same method to simulate these flips except we will
save new objects to our R environment. <br />



```{r}
bigSim <- sample(x, size = 100000, replace = TRUE)
bigLogic <- (bigSim =="H" )
howBig <- length(which(bigLogic))


```

Compare how many heads you got by calling length() on the object where you stored
only heads in our outcome (bigLogic) with how many times the coin was flipped.

```{r}
length(which(bigLogic))
```


In our simultion, `r howBig` heads came up. This outcome is very
close to 50,000 which is the outcome we would expect if the coin
were fair and we flipped it 100,000 times like we did. 



## If you want to use a variable in your text, like a p value

You want to add a echo=false snippet and call it in the text
using an embedded feature. 

This is very, very powerful.


See below: 

```{r, echo = FALSE}

x <- sample(1:1000, 100000, replace = TRUE)
logic <- (x < 100)
how_many <- length(which(logic))

```

There are `r how_many` elements of $x$ which are $<100$







