---
title: 'AMOD 5240H - Workshop #07'
author: "Wesley Burr"
date: "November 14, 2017"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Difference in Means - Simulation

Today we are going to examine two datasets from the "Lock5" textbook. They are contained in an R package called **Lock5withR**. Install it first.

```{r}
# install.packages("Lock5withR")
library("Lock5withR")
```

We are interested in the problem of differences of means. We are going to use the American Community Survey (ACS) - Employment Data. 

```{r}
data("EmployedACS")
str(EmployedACS)
male_income <- EmployedACS[EmployedACS$Sex == 1, "Income"]
female_income <- EmployedACS[EmployedACS$Sex == 0, "Income"]
mean(male_income)
mean(female_income)
```

Now, let's compute a point estimate of the difference of mean:
```{r}
point_est <- mean(male_income) - mean(female_income)
SE <- sqrt( var(male_income)/length(male_income) + var(female_income)/length(female_income) )
test_stat0 <- point_est / SE
test_stat0
```

Now, we would like to use permutation methods and create a simulation for testing this. We need to take all of the incomes, and sample for the men and the women. This is one permutation:

```{r}
total_income <- c(male_income, female_income)
n_men <- length(male_income)
n_fem <- length(female_income)
samp <- sample(total_income, size = n_men + n_fem, replace = FALSE)
samp_men <- (samp[1:n_men])
samp_fem <- (samp[(n_men + 1):(n_men + n_fem)])
test_stat <- (mean(samp_men) - mean(samp_fem)) / sqrt( var(samp_men)/n_men + var(samp_fem)/n_fem)
```

Now, let's turn this into a proper simulation. 
```{r}
res <- vector(length = 10000)
for(j in 1:10000) {
  samp <- sample(total_income, size = n_men + n_fem, replace = FALSE)
  samp_men <- (samp[1:n_men])
  samp_fem <- (samp[(n_men + 1):(n_men + n_fem)])
  res[j] <- (mean(samp_men) - mean(samp_fem)) / sqrt( var(samp_men)/n_men + var(samp_fem)/n_fem )
}                                                      
hist(res)
abline(v = test_stat0, col = "blue")
```

Let's formally state this:
$H_0:$ $\mu_1 - \mu_2 = 0$ versus $H_A$: $\mu_1 > \mu_2$

Then, our empirical *p*-value is computed as:
```{r}
p_val <- length( which ( res >= test_stat0 ) ) / 10000
p_val
```

The conclusion is then that we reject the null hypothesis, and conclude that the data provide convincing evidence of $\mu_1 > \mu_2$, in other words, male income is higher than female income for America in 2011.

Let's conclude by doing this the "fast" way. 
```{r}
1 - pnorm(test_stat0)
```

## Computing ANOVA in R

You can do ANOVA by hand, but it's rather difficult. In this section, we'll explore an ANOVA problem in R, starting from a raw data set. This first example is of an ANOVA for regression (total model fit), something we'll talk about tomorrow in class.

We are interested in whether the relationship between Calories and Sugars.
```{r}
library("Lock5withR")
data(Cereal)
plot(x = Cereal$Sugars, y = Cereal$Calories, type = "p")
fit <- lm(Calories ~ Sugars, data = Cereal)
summary(fit)
aov(fit)
```

The key is setting up a properly formatted data.frame object. If it's in a form ready for regression, it's ready for you to use.

If you want to compute an ANOVA on a multiple-factor problem, like the questions on Assignment 3, it's slightly more complicated. We'll use a data set on Fish Gills, again from the Lock5 data sets.

```{r}
data(FishGills3)
head(FishGills3)
levels(FishGills3$Calcium)
```

As you can see from this data set, FishGills3 contains two columns: Calcium and GillRate. There are three levels of Calcium: Low, Medium, and High. This means we can split the data into three groups, based on Calcium levels. 

We need to examine the data. The ggplot-family package ggpubr is great for this.

```{r}
install.packages("ggpubr", repos = "https://cloud.r-project.org")
library("ggpubr")
ggboxplot(FishGills3, x = "Calcium", y = "GillRate", 
          color = "Calcium", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("Low", "Medium", "High"),
          ylab = "Gill Rate", xlab = "Calcium Levels")
```

The variances of the three groups appear to be the same. Examine the means:

```{r}
ggline(FishGills3, x = "Calcium", y = "GillRate", 
       add = c("mean_se", "jitter"), 
       order = c("Low", "Mean", "High"),
       ylab = "Gill Rate", xlab = "Calcium Levels")
```

Let's proceed with our ANOVA. It's aov() again, just with a Factor variable established.

```{r}
# Compute the analysis of variance
res.aov <- aov(GillRate ~ Calcium, data = FishGills3)
# Summary of the analysis
summary(res.aov)
```

To access the results, notice that the summary(res.aov) is actually a list of one element containing a data.frame (more list elements):
```{r}
str(summary(res.aov))
```

which means we can access the elements by doing:

```{r}
tab_anova <- summary(res.aov)[[1]]
str(tab_anova)
```

Then all the elements are list elements:

```{r}
tab_anova[[1]]
tab_anova[[2]][1]
```

So to get the p-value extracted, we could use
```{r}
tab_anova[[5]][1]
```

This concludes our example.