---
title: "Workshop 4 - visualization"
author: "Nicholas Hopewell"
date: "October 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic plots


LOOOK UP ALL THE ARGUEMNTS FOR ?PLOT 

All we need is Y..

```{r}

y <- rnorm(50, mean = 10, sd =2)
plot(y)

```


rnorm creates random normal variables

This is not very pretty. We can tweak it. 

```{r}
x <- paste0("A", 1:50)
plot(1:50, y, xaxt = 'n')
axis(side = 1, at = 1:50, labels = x)

```
The above code makes the axis not appear and then puts it in manually. 
This is good if u need to change the measurement of the axis. 


Another example where we subset:

```{r}
plot(1:50, y, xaxt = 'n')
axis(side =1, at = seq(1:50, 4), labels= x[seq(1:50, 4)])

```


## Plot arguemnts 

Lets start with a fully expanded plot to demo some arguments
Here is a bad example where the variable is going off the screen:
```{r}
x <- 1:20
y <- rnorm(20)
plot(x,y, type = "l", col = "red", xlab = "X var", ylab = "YVar", main = "My Plot's Title")
y2<- rnorm(20) +0.3
lines(x,y2,col = "blue")

```


To fix this so the lines all fit we compute the min and the max before we start:

```{r}
x <- 1:20
y <- rnorm(20)
ymax <- max(y, y2)
ymin <- min(y, y2)
plot(x,y, type = "l", col = "red", xlab = "X var", ylab = "YVar", main = "My Plot's Title", ylim = c(ymin, ymax))
y2<- rnorm(20) +0.3
lines(x,y2,col = "blue")

```


But which line is which? Add a legend:
add points as well not as a line just as points then add that to legend
Notice how we ask the legend to fill with colour since we added dots and we cant covery that as a black
line in the legend
```{r}
x <- 1:20
y <- rnorm(20)
ymax <- max(y, y2)
ymin <- min(y, y2)
plot(x,y, type = "l", col = "red", xlab = "X var", ylab = "YVar", main = "My Plot's Title", ylim = c(ymin, ymax))
y2<- rnorm(20) +0.3
lines(x,y2,col = "blue")
points(x, rnorm(20), col = "black")
legend(x = "bottomright", legend = c("Red Line", "Blue Line"), 
      col = c("red", "blue"), lty = c(1,3),
      fill = c("red", "blue", "black"))



```



## Types of plots" the *type argument


the argument **type**. we have used it to create **l** (lines) and **p** (points)

Type = 

**p** points
**l** lines
**b** both lines and points
**c** lines wiht holes where the poins should be
**h** bar-plot / histogram like display

display graphs in the same frame (par function), displaying the same plot in 4 ways:

```{r}

x <- 1:15
y <- rnorm(15,10,3)
plot(x,y, type = "l")
plot(x,y, type = "b")
plot(x,y, type = "c")
plot(x,y, type = "h")

```


The histrgram like plot looks funky, lets chang the bar width. 
Also change the end of the lines to squared off lines and not rounded (lend = 1)
```{r}
x <- 1:15
y <- rnorm(15,10,3)
plot(x,y, type = "h", lwd = 15, lend = 1, col = "grey60")

```

## And now for something completely different...

**data frames** are objects which store data in R, where the storage is done by column, and each column can have its own data type - that last part is the most important part. When you make a matrix they are all the same types of data, in data frames you can have any combination of data types. 

The call to creat a dataframe is data.frame...

```{r}

t <- data.frame(X =c(1,2,3),
                Y =c("a", "b", "c"),
                Z =c(TRUE, TRUE, FALSE))

str(t)


# R will automatically convert string data types to factors,
#to stop r from converting Y to a factor, indicate that in the last line.
# this is very very important. If you specify it, r wont change ur data
u <- data.frame(X =c(1,2,3),
                Y =c("a", "b", "c"),
                Z =c(TRUE, TRUE, FALSE),
                stringsAsFactors = FALSE)
```


Tomorrow we will demonstrat reading data from files, and obtaining data.frame


**ALWAYS** read in your data and call str() on it to make sure it is what you need it to be because it usually will not be. 


## Access of a Data frame sub-objects

Tje \$ operator in R accesses sub:objects of objects **by name**. If the object does not have named sub-objects.


There are three essential operators for accessing data in R. they are: 

* [[
* [
* $


data frames are actually vectorized pointers to list, each element is actually a list. 
The [[]]  operator says give me a certain element of the list.
```{r}
cc

u[[2]]
u$Y
u[["Y"]]
```

What happens if you need to pull things out by name, and we dont know the names?

```{r}


u <- data.frame(X =c(1,2,3),
                Y =c("a", "b", "c"),
                Z =c(TRUE, TRUE, FALSE),
                stringsAsFactors = FALSE)

names(u)
u[[names(u)[2]]]


```

What happens if we create a data.frame by casting a matrix?

```{r}
x <- matrix(data = 1:12, nrow= 4, ncol = 3)
y <- as.data.frame(x)
names(y) <- c("X", "Y", "Z")


```
