---
title: "Workshop 5 - reading data/chi squared"
author: "Nicholas Hopewell"
date: "October 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

SEE HIS FILE FOR THE FIRSTPART 

Compute chi-squared test statistic for independence 

cont_table in his example were the observed values
expected = the expected values

```{r}

# get the sum of the difference between observed and expected squared divided by expected
chisq_test <- sum((cont_table - expected)^2 / expected)


# get the degrees of freedom:
df <- (2-1) * (3-1)  #rows -1 * columns - 1

# get p-values

p_val <- pchisq(chisq_test, df = df, lower.tail = FALSE)

```


So, for a NHST of independence, a *p*-value of `r p_val` means that we do not have evidence at a $\alpha=0.05$ level to reject the null hypothesis and conclude that the two variables are dependent. 



## Makin functions

see the part of his code where he found the row and column totals as well as the overall total? It was very lengthy. lets do it in a function

```{r}

# he saved the function as a script file then sourced it before calling the function
# there should be 2 files on blackboard

chisq_test <- function(observed) {
  
  
  
  
}


```